Morphological diversity in tenrecs (Afrosoricida, Tenrecidae): Comparing tenrec skull diversity to their closest relatives

### 2015-01-28 paper/Supplemental.tex
Supplemental for PeerJ submission

\subsubsection{Skulls: dorsal view}
\vspace*{-0.5cm}
	Most of our landmarks in this view are relative (type 3) points which represent overall morphological shape but not necessarily homologous biological features \citep{Zelditch2012}. We placed ten landmarks and drew four semilandmark curves to represent the shape of both the braincase (posterior) and nasal (anterior) area of the skulls. Table \ref{tab:skdors} describes how we defined the landmarks and outline curves for our images of skulls in dorsal view.

%Table with skdors landmark descriptions
\begin{table}[h]
	\caption[Skulls: dorsal landmarks]
		{Descriptions of the landmarks (points) and curves (semilandmarks) for the skulls in dorsal view} 
	\input{tables/skdors.landmarks}
	\label{tab:skdors}
\end{table}
%--------------------------------------------------------------

\subsubsection{Skulls: ventral view}
\vspace*{-0.5cm}
	Most of the landmarks in this view are concentrated around the dentition and palate of the animals. We identified 13 landmarks and drew one outline curve (resampled to 60 semilandmark points) around the back of the skull between landmarks 12 and 13. The high variability of our species' basi-cranial regions and difficulties associated with identifying developmentally or functionally homologous points precluded designation of additional landmarks towards the back of the skulls. Table \ref{tab:skvent} outlines the descriptions of the landmarks we used for the ventral skull photos.

% Skulls ventral landmarks
\begin{table}[!htb] 
\caption[Skulls: ventral landmarks]
		{Descriptions of the landmarks (points) and curves (semilandmarks) for the skulls in ventral view.} 
\input{tables/skvent.landmarks}
\label{tab:skvent}
\\
*This curve does not necessarily trace homologous features because of the variation in the position of the foramen magnum.
\end{table}

%------------------------------------------------------
\newpage
\subsubsection{Skulls: lateral view}
\vspace*{-0.5cm}
	We placed nine landmarks on the photographs of skulls in lateral view and also drew two semilandmark curves to represent the shape of the back of the skull (between landmarks 7 and 8, re-sampled to 20 semilandmark points) and the top of the skull (between landmarks 8 and 1, re-sampled to 15 semilandmark points). Table \ref{tab:sklat} describes our definitions for each of the landmark points.
	If specimens were damaged on their right side we reflected photographs of the left lateral side of the skull so that all the images were in the same orientation.

% Skulls lateral landmarks
\begin{table}[!htb]
\caption[Skulls: lateral landmarks]
		{Descriptions of the landmarks (points) and curves (semilandmarks) for the skulls in lateral view.} 
\input{tables/sklat.landmarks}
\label{tab:sklat}
\end{table}
%-------------------------------------------
\subsection{{Potential morphometrics errors}}
\vspace*{-0.5cm}	 
	While 2D methods are an accepted means of comparing morphological shape \citep[e.g.][]{Adams2004, Mitteroecker2009}, particularly for comparing skull morphologies of small mammals \citep[e.g.][]{Cardini2003, Panchetti2008, White2008, Barrow2008, Scalici2011}, the inherent discrepancies associated with comparing three dimensional objects using two dimensional pictures can introduce some potential problems of possible image distortion \citep{Arnqvist1998}. Similarly, human error with how landmarks are positioned on specimens could also introduce noise into further analyses.
	In contrast to detailed intraspecific work \citep[e.g.][]{Bornholdt2008, Blagojevic2011}, photographic or landmark placement errors are unlikely to be significant in our interspecific study since one would expect that the morphological variation among species is large enough to  be detected as a signal above any background noise associated with methodological error \citep{Arnqvist1998}. Nevertheless, it is still important to assess measurement error in a morphometric data set to increase confidence in the outcome of final analyses.
	We identified two possible sources of morphometric measurement error: specimen orientation and placement of landmarks.

@@

>>>>

\subsubsection{Skulls: dorsal view}
\vspace*{-0.5cm}
	Most of our landmarks in this view are relative (type 3) points which represent overall morphological shape but not necessarily homologous biological features \citep{Zelditch2012}. We placed ten landmarks and drew four semilandmark curves to represent the shape of both the braincase (posterior) and nasal (anterior) area of the skulls. Table S1 describes how we defined the landmarks and outline curves for our images of skulls in dorsal view.
%--------------------------------------------------------------

\subsubsection{Skulls: ventral view}
\vspace*{-0.5cm}
	Most of the landmarks in this view are concentrated around the dentition and palate of the animals. We identified 13 landmarks and drew one outline curve (resampled to 60 semilandmark points) around the back of the skull between landmarks 12 and 13. The high variability of our species' basi-cranial regions and difficulties associated with identifying developmentally or functionally homologous points precluded designation of additional landmarks towards the back of the skulls. Table S2 outlines the descriptions of the landmarks we used for the ventral skull photos.

%------------------------------------------------------
\newpage
\subsubsection{Skulls: lateral view}
\vspace*{-0.5cm}
	We placed nine landmarks on the photographs of skulls in lateral view and also drew two semilandmark curves to represent the shape of the back of the skull (between landmarks 7 and 8, re-sampled to 20 semilandmark points) and the top of the skull (between landmarks 8 and 1, re-sampled to 15 semilandmark points). Table S3 describes our definitions for each of the landmark points.
	If specimens were damaged on their right side we reflected photographs of the left lateral side of the skull so that all the images were in the same orientation.

%-------------------------------------------
\subsection{{Potential morphometrics errors}}
\vspace*{-0.5cm}	 
	While 2D methods are an accepted means of comparing morphological shape \citep[e.g.][]{Adams2004, Mitteroecker2009}, particularly for comparing skull morphologies of small mammals \citep[e.g.][]{Cardini2003, Panchetti2008, White2008, Barrow2008, Scalici2011}, the inherent discrepancies associated with comparing three dimensional objects using two dimensional pictures can introduce some potential problems of possible image distortion \citep{Arnqvist1998}. Similarly, human error with how landmarks are positioned on specimens could also introduce noise into further analyses.
	In contrast to detailed intraspecific work \citep[e.g.][]{Bornholdt2008, Blagojevic2011}, photographic or landmark placement errors are unlikely to be significant in our interspecific study since one would expect that the morphological variation among species is large enough to  be detected as a signal above any background noise associated with methodological error \citep{Arnqvist1998}. Nevertheless, it is still important to assess measurement error in a morphometric data set to increase confidence in the outcome of final analyses.
	We identified two possible sources of morphometric measurement error: specimen orientation and placement of landmarks.

@@

%--------------------------------------------
%Table legends
%Table S1 Skulls: dorsal landmarks
%Descriptions of the landmarks (points) and curves (semilandmarks) for the skulls in dorsal view.

% Table S2 Skulls: ventral landmarks
%Descriptions of the landmarks (points) and curves (semilandmarks) for the skulls in ventral view.

% Table S3 Skulls: lateral landmarks
%Descriptions of the landmarks (points) and curves (semilandmarks) for the skulls in lateral view.
